---
title: "Assigment1"
author: "Ernesto"
date: "10/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First un zip and load the data




```{r}
if (!file.exists('activity.csv')) {
  unzip(zipfile = "activity.zip")
}

activity <- read.csv(file="activity.csv", header=TRUE)
```

# What is mean total number of steps taken per day? 

Calculate and plot histogram

```{r}
# Calculate the total steps taken per day
daysteps <- aggregate(steps ~ date, activity, FUN=sum)

# Make a histogram 
hist(daysteps$steps,
     main = "Steps per Day",
     xlab = "Steps")
```

the mean and median total number of steps taken per day
```{r}
meansteps <- mean(daysteps$steps, na.rm = TRUE)
mediansteps <- median(daysteps$steps, na.rm = TRUE)
print(meansteps)
print(mediansteps)

```
##What is the average daily activity pattern?

plot the time series
```{r}
# Calculate the mean for interval
intersteps <- aggregate(steps ~ interval, activity, FUN=mean)
plot.ts(intersteps$steps)

```

Max interval 

```{r}
maxint <- intersteps[which.max(intersteps$steps),]
print(maxint)
```

##Imputing Missing Values
```{r}
# Calculate and report the total number of missing values
missingVals <- is.na(activity$steps)
print(sum(missingVals))
```
imputing NA with the average value
```{r}
new_activity <- transform(activity,
                          steps = ifelse(is.na(activity$steps),
                                        intersteps$steps[match(activity$interval, 
                                                               intersteps$interval)],
                                             activity$steps))
head(new_activity)
```
Make histogram.

```{r}
# Make a histogram of the total number of steps taken each day
new_daystep <- aggregate(steps ~ date, new_activity, FUN=sum)
hist(new_daystep$steps,
     main = "Imputed Number of Steps Per Day",
     xlab = "Steps")
```
Reporting the mean and median
```{r}
new_mean <- mean(new_daystep$steps, na.rm = TRUE)
new_median <- median(new_daystep$steps, na.rm = TRUE)
print(diffmean <- new_mean - meansteps)
print(diffmed <- new_median - mediansteps)
print(difftotal <- sum(new_daystep$steps) - sum(daysteps$steps))
```
##Are there differences in activity patterns between weekdays and weekends?

Make a factor variable
```{r}
# Create a new factor variable with two levels - "weekend" and "weekday"
DayType <- function(date) {
  day <- weekdays(date,abbreviate = FALSE)
  if (day %in% c('lunes', 'martes', 'miércoles', 'jueves', 'viernes'))
      return ("weekeday")
  else if (day %in% c('sábado', 'domingo'))
      return ("weekend")
  else
       print(weekdays(date))
}
print(missingdates <- sum(is.na(activity$date)))
new_activity$date <- as.Date(new_activity$date)
new_activity$day <- sapply(new_activity$date, FUN = DayType)
```
plotting time series

```{r}
library(ggplot2)
weekstapes <- aggregate(steps ~ interval + day, new_activity, mean)
ggplot(data = weekstapes, aes(x = interval, y = steps)) + 
  geom_line() +
  facet_grid(day ~ .) +
  ggtitle("Average Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average of Steps") +
  theme(plot.title = element_text(hjust = 0.5))
```